---
title: "Lab 3"
date: last-modified
format: 
  html:
    code-overflow: wrap
---

# General Instructions

- Submit your work via [Canvas](https://canvas.case.edu/). 
- The deadline for this Lab is specified on the [Calendar](calendar.qmd).
  - Work submitted more than 59 minutes late, but within 12 hours of the deadline will lose 5 of the available 50 points.
  - Work submitted 12 to 24 hours after the deadline will lose 10 of the available 50 points.
  - Work submitted more than 24 hours after the deadline will not be graded.

Your response should include a Quarto file (.qmd) and an HTML document that is the result of applying your Quarto file to the data we've provided. While we have not provided a specific template for this Lab, we encourage you to adapt the one provided for [Lab 2](lab2.qmd#template).

## The Data

This Lab uses the `hbp3456` data developed in [Lab 2](lab2.qmd). See the [Lab 2 instructions](lab2.qmd) for details on the data set. Back in Lab 2, we loaded the data with this code.

```{r}
#| message: false
#| warning: false

library(janitor)
library(tidyverse)

knitr::opts_chunk$set(comment = NA)

hbp3456 <- read_csv("https://raw.githubusercontent.com/THOMASELOVE/432-data/master/data/hbp3456.csv", 
                    show_col_types = FALSE) |>
  clean_names() |>
  mutate(record = as.character(record))
```

Here, we will walk through the process of fitting and evaluating linear regression fits to predict a subject's estimated (neighborhood) median income (the `income` variable) on the basis of the following five predictors:

- the subject's neighborhood high school graduation rate, collected in the `hsgrad` variable
- the subject's race category, from the `race` variable
- the subject's Hispanic/Latinx ethnicity category, as shown in `eth_hisp`,
- the subject's age (in the `age` variable), and 
- the subject's current tobacco status, available in the `tobacco` variable. 

## Preliminary Data Work for Lab 3

Start your work by completing the following tasks to create a tibble that we'll call `hbp_b` in the answer sketch:

1. Exclude the 25 subjects in `hbp3456` who have missing values of either `hsgrad` or `income`.
2. Restrict your data to the variables we'll use in our models (the five predictors listed above, the estimated neighborhood income, and the subject identifying code (the `record`)).
3. Ensure that all character variables (other than `record`) in your tibble are recognized as factors.
4. Create a new variable called `sqrtinc` which will serve as your response (outcome) for your regression modeling, within your tibble.
5. Use `set.seed(432)` and `slice_sample()` to select a random sample of 1000 subjects from the tibble.

Your resulting `hbp_b` tibble should look like this:

```{r}
#| echo: false
#| message: false

hbp_b <- read_csv("https://raw.githubusercontent.com/THOMASELOVE/432-data/master/data/hbp3456.csv", show_col_types = FALSE) |>
    filter(complete.cases(hsgrad, income)) |>
    select(record, income, hsgrad, race, eth_hisp, age, tobacco) |>
    type.convert(as.is = FALSE) |>
    mutate(record = as.character(record), sqrtinc = sqrt(income))

set.seed(432)
hbp_b <- hbp_b |>
    slice_sample(n = 1000)
```

```{r}
hbp_b
```

**Note** If your tibble looks different, it's not immediately clear to me why. In the answer sketch, I do the following things, in this order:

1. read in the data (note that whether you then clean_names() or not has no real effect in this particular case)
2. filter to complete cases on hsgrad and income
3. select the seven variables we'll use
4. convert all character variables to factors
5. ensure record remains a character variable
6. add the `sqrtinc` variable to the tibble
7. set a seed of 432
8. create the random sample

# Question 1. (10 points)

Display the code you used to ingest the data and complete the preliminary data work described above. Be sure to include text annotations to clarify exactly what your code is doing. Then produce a table to tell us how many missing values you have in each of the important variables in your `hbp_b` tibble. The important variables are your outcome (square root of estimated neighborhood income) and the five predictors.

# Question 2. (10 points)

Using the entire sample in `hbp_b`, obtain and display an appropriate Spearman $\rho^2$ plot and use it to identify a good choice of a single non-linear term that adds exactly two degrees of freedom to the main effects model using all five predictors for `sqrtinc`. Specify your choice of non-linear term, and your motivation for that choice, based on the plot.

# Question 3. (10 points)

Fit the main effects model for `sqrtinc` using `ols` in the `hbp_b` sample, and call that model `m1`. Plot the effect summary (using `plot(summary(m1))`) for model `m1` and carefully explain the meaning of the `hsgrad` coefficient shown in that plot in a complete English sentence. 

- Hint 1: you are permitted to also fit the model using `lm`, if that is useful to you.
- Hint 2: If you use `anova()` on model `m1` you should have 8 total degrees of freedom in your model.

# Question 4. (10 points)

Fit a new model using `ols`, for `sqrtinc` using all five predictors, including the non-linear term you identified in Question 2 in the `hbp_b` sample, and call that model `m2`. Plot the effect summary (using `plot(summary(m2))`) for model `m2`, and explain the meaning of the `tobacco` coefficient shown in the plot in a complete English sentence.

- Hint 1: you are permitted to also fit the model using `lm`, if that is useful to you.
- Hint 2: If you use `anova()` on model `m2` you should have 2 non-linear degrees of freedom, and 10 total degrees of freedom in your model.

# Question 5. (10 points)

You've now fit models `m1` and `m2`. For each model, obtain the following summary statistics: the uncorrected raw $R^2$ value, the AIC and BIC. Then validate each model's $R^2$ and MSE values using `set.seed(2023)` and 40 bootstrap replications. 

Now, report the five results you obtained for each model in an attractive, well-formatted table. Then write a sentence or two explaining what your findings mean about the performance of the two models.

# Our Best Advice

Review your HTML output file carefully before submission for copy-editing issues (spelling, grammar and syntax.) Even with spell-check in RStudio (just hit F7), it's hard to find errors with these issues in your Quarto file so long as it is running. You really need to look closely at the resulting HTML output.

# Use of AI

If you decide to use some sort of AI to help you with this Lab, we ask that you place a note to that effect, describing what you used and how you used it, as a separate section called "Use of AI", after your answers to our questions, and just *before* your presentation of the Session Information. Thank you.

# Session Information

Please display your session information at the end of your submission, as shown below.

```{r}
xfun::session_info()
```

## After the Lab

We will post an answer sketch 24 hours after the Lab is due.

We will post grades to our Grading Roster on our Shared Google Drive one week after the Lab is due.

See the [Lab Appeal Policy in our Syllabus](https://thomaselove.github.io/432-syllabus-2024/08-grading.html#lab-appeal-policy---request-a-review-via-google-form) if you are interested in having your Lab grade reviewed, and use [the Lab Regrade Request form specified there](https://bit.ly/432-2024-lab-regrades) to complete the task. Thank you.